---
title: "A quick look at GDP components over time"
author: "Group 5: Etra Bianco, Himanshu Singh, Yi Yan Ng, Mengze Sun, Christian Gonder, Vasu Dev Puri"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
library(knitr)
library(kableExtra)
```


The GDP data we will look at is from the [United Nations' National Accounts Main Aggregates Database](https://unstats.un.org/unsd/snaama/Downloads), which contains estimates of total GDP and its components for all countries from 1970 to today. We will look at how GDP and its components have changed over time, and compare different countries and how much each component contributes to that country's GDP. At the risk of oversimplifying things, the main components of gross domestic product, GDP are personal consumption (C), business investment (I), government spending (G) and net exports (exports - imports). 


```{r read_GDP_data,cache=TRUE}

UN_GDP_data  <-  read_excel(here::here("/content/projects/gdpcountries/data", "Download-GDPconstant-USD-countries.xls"), # Excel filename
                sheet="Download-GDPconstant-USD-countr", # Sheet name
                skip=2) # Number of rows to skip

```


We first tidy the data and express all figures in billions.

```{r}

tidy_GDP_data  <- UN_GDP_data  %>% 
  pivot_longer(names_to = 'year', values_to  = 'Dollar Amount', cols = -c('CountryID', 'Country', 'IndicatorName')) 
tidy_GDP_data$`Dollar Amount` <- tidy_GDP_data$`Dollar Amount` / 1e9


tidy_GDP_data <- tidy_GDP_data %>%
  pivot_wider(names_from = IndicatorName, values_from = `Dollar Amount`) %>%
   rename("Household expenditure" = "Household consumption expenditure (including Non-profit institutions serving households)", "Government expenditure" = "General government final consumption expenditure", "Exports" = "Exports of goods and services", "Imports" = "Imports of goods and services")
  
# Let us compare GDP components for these 3 countries
country_list <- c("United States","India", "Germany")
```


We choose a few countries to explore the data first. Here we use Germany, India and the United States to plot a graph of their GDP over the years.

```{r, fig.width=10,fig.asp=0.6}

data_for_plot  <- tidy_GDP_data    %>% 
  select(1:3, 5:7,9 ,10) %>%
  filter(Country %in% c("Germany", "India", "United States")) %>%
  pivot_longer(cols = 4:8, names_to = "GDP component", values_to = "Dollar Amount")

data_for_plot

plot <- data_for_plot %>%
ggplot() + 
         geom_line(aes(x=as.Date(paste0("02-02-",`year`),"%d-%m-%Y"), y=`Dollar Amount`, color=`GDP component`)) + 
  facet_wrap(~Country) + 
  labs(title = "GDP components over time", subtitle = "In constant 2010 USD", color = "Components of GDP", x = "", y = "Billion US$") +
  theme_bw()

plot
         
  
```

From the chart above we can see that for all 3 countries Net Exports represents the smallest component of GDP, and vice versa, Household expenditure is the largest but its worthwhile that the latter has evolved differently in time across these countries (rising in the US, decreasing in India and plateauing in Germany). Gross capital formation has been rising sharply in India since 2000, whilst it has remained relatively constant in the US and Germany. Government expenditure instead has been fairly constant across all 3 countries, with a slight decline for the US and it has the same weight on GDP as Gross Capital formation for US and Germany, whilst it is much smaller component for India. 


Next, we note that GDP is the sum of Household Expenditure (Consumption *C*), Gross Capital Formation (business investment *I*), Government Expenditure (G) and Net Exports (exports - imports) and calculate it as follows. 
```{r}
options(scipen = 999)

calculated_GDP <- tidy_GDP_data %>%
mutate(Net_Exports = `Exports`-`Imports`) %>%
# *******double check what columns should be included in the TotGDP count column below**********
  mutate(TotGDP = `Household expenditure`+ `Government expenditure`+ `Net_Exports` + `Gross capital formation`) %>%
  select(2,3,11, 22) %>%
  mutate(Percentage_Difference = (`TotGDP`-`Gross Domestic Product (GDP)`)/`Gross Domestic Product (GDP)`*100)


  
calculated_GDP

GDP_calculations_for_2017 <- calculated_GDP %>%
  filter(year == 2017)

GDP_calculations_for_2017 %>%
summarize(mean = mean(!is.na(Percentage_Difference)))

GDP_Difference <- calculated_GDP %>%
  filter(year == 2017, Country %in% c("Italy", "Afghanistan", "Germany", "India","Mozambique"))

GDP_Difference 



```
The percentage difference between what we calculated as GDP and the GDP figure included in the dataframe varies from country to country and from year to year. However, we observe that the average difference in 2017 was, on average, less than 1%. When looking more closely at a handful of countries it becomes evident that there are exceptions, such as with Afghanistan, which has a % difference of 31%. 


Plotting the same graph for my home country Singapore, we can see that the imports and exports represent the largest component of the GDP in Singapore, which has been on a sharp increase since 1990s. In comparison, the rest of the components have remained relatively constant.
```{r,fig.width=10,fig.asp=0.6}

data_for_plot  <- tidy_GDP_data    %>% 
  select(1:3, 5:7,9 ,10) %>%
  filter(Country %in% c( "Singapore")) %>%
  pivot_longer(cols = 4:8, names_to = "GDP component", values_to = "Dollar Amount")

data_for_plot

plot <- data_for_plot %>%
ggplot() + 
  geom_line(aes(x=as.Date(paste0("02-02-",`year`),"%d-%m-%Y"), y=`Dollar Amount`, color=`GDP component`)) + 
  facet_wrap(~Country) + 
  labs(title = "GDP components over time", subtitle = "In constant 2010 USD", color = "Components of GDP", x = "", y = "Billion US$") +
  theme_bw()

plot
         
  
```


